### 2024-08-31 article_codes.qmd

---
execute:
  warning: false
  message: false
---

# R scripts

```{r}
#| label: libraries

# Libraries
library(tidyverse)
library(terra)
library(tidyterra)
library(sf)
library(patchwork)
library(gt)
library(gtExtras)
library(purrr)

terraOptions(progress = 0)

```

## Functions

In addition to the packages, we define our own functions for the analysis:

```{r}
#| label: aux-functions

# Function: get_perc_veg_buffer ----
#' Function for calculating if the percentage of vegetation within a buffer area is higher than a threshold 
#' @description It returns a raster indicating if the percentage of flammable vegetation within a circumference, from the centre of the grid (buffer), higher than a threshold (YES = 1, NO = 0)
#' @param .veg raster layer indicating if the vegetation is considered as fuel or not (YES = 1, NO = 0). 
#' @param .radio_buffer Radio of the circumference.
#' @param .threshold Proportion: from 0-1 (default  = 0.5; i.e., 50%) 
get_perc_veg_buffer <- function(.veg,
                                .radio_buffer,
                                .threshold = 0.5){

  # Percentage-vegetation 
  fwm <- terra::focalMat(.veg, d = radio_buffer, type = "circle", fillNA = TRUE)
  focal_perc <- terra::focal(.veg, w = fwm, fun = "sum", na.rm = TRUE)

  # Higher than threshold (YES = 1, No = 0)
  wv_perc <- focal_perc
  wv_perc[wv_perc <  .threshold] <- 0
  wv_perc[wv_perc >= .threshold] <- 1

  return(wv_perc)

}

# Function: get_large_polygons ----
#' Function for removing small areas from a raster file. Used for calculating large forest areas (e.g., > 5 km2).  
#' @description It returns a raster with all the polygons larger than a minimal area (YES = 1, NO = 0)
#' @param .x starts object (i.e., 1 - fuel vegetation, 0 - no fuel vegetation). 
#' @param .min_area Minimum area [in km2] for considering an vegetated area as large  (Default 5 km^2)
#' @param .resolution Resolution of the final raster (default - 100 x 100 m)
get_large_polygons <- function(.x,               
                               .min_area = 5,    
                               .resolution = 100) {

  df_v <- .x |> 
    # Convert to polygons
    st_as_sf(as_points = FALSE, merge = TRUE) %>% 
    # Calculate the area of the polygons
    mutate(area = units::set_units(st_area(.), "km2")) |> 
    # Remove lower than 5 km2
    filter(area > units::set_units(.min_area, "km2")) |> 
    # Set columns names
    setNames(c("large_wild_veg", "geometry", "area")) |> 
    # Remove area column
    select(-area) |> 
    # transform to SpatVector  
    vect() |> 
    # aggregate in one polygon
    aggregate(by = "large_wild_veg") |> 
    select(large_wild_veg)

  # Rasterize
  df_geom <- rast(df_v, resolution = .resolution)
  df_r <- rasterize(df_v, df_geom)

  return(df_r)

}

# get_dist_max ----
#' Get cells within a distance to vegetated area (YES --> 1; NO --> 0)  
#' @description It returns a raster with indicating the cells that are within the distance to the vegetated areas (YES = 1, NO = 0)
#' @param .x Raster with distances (from `terra::distance`) 
#' @param .max_distance Maximum distance in [m] (e.g., 2400 m)
get_dist_max <- function(.x, .max_dist) {

  dist <- .x
  dist[dist <= .max_dist] <- 1
  dist[dist >  .max_dist] <- 0

  return(dist)

}

# Function: get_wui_raster ----
#' Function for WUI classification
#' @description it returns a raster with WUI areas (Interface or Intermix)
#' @param .veg_perc raster layer indicating if the percentage of vegetation in
#' the buffer area is higher than a threshold (e.g., > 50%). (YES = 1, NO = 0). 
#' @param .veg_dist raster layer indicating if the a grid cell is within a 
#' distance (e.g., 2.4 km) of a large area of flammable vegetation (YES = 1, NO = 0)
#' @param .pot_wui raster layer with potential WUI grids based on building density (`wui_potential`).
#' @param .region SpatVector layer with the region area (`region_v`) 
get_wui_raster <- function(.veg_perc,
                           .veg_dist,
                           .pot_wui = wui_potential,
                           .region = region_v){

  # Potential WUI grid ----
  wui_potential <- .pot_wui
  wui_potential <- as.factor(wui_potential)

  # Match extents
  perc <- resample(.veg_perc, wui_potential, method = "mode")
  dist <- resample(.veg_dist, wui_potential, method = "mode")

  # Intermix 
  wui <- perc + wui_potential # 2 = WUI intermix
  wui[wui == 1] <- 0          # 1 = Potential interface --> 0

  # Interface 
  wui <- wui + dist
  wui[wui == 0] <- NA # No WUI
  wui[wui == 1] <- 1  # Interface
  wui[wui == 2] <- 2  # Intermix
  wui[wui == 3] <- 2  # Intermix

  # NAs as 0 inside the study area (otherwise NA)
  wui <- ifel(is.na(wui), 0, wui)

  # Mask to region (e.g., NA = ourside norway)
  wui <- terra::mask(wui, .region)

  # Identify levels (1 = Interface, 2 = Intermix)
  levels(wui) <- data.frame(id = 0:2,
                            wui_class = c("No_wui", "Interface", "Intermix")
                            )

  # Add colour table 
  coltab(wui) <- data.frame(id = 0:2, 
                            col = c("#FAFAFA", "#FF6633", "#003366"))

  # Return SpatRaster
  set.values(wui)
  return(wui)

}

# add_colours_wui 
#' Function for adding WUI colours in a SpatRaster 
#' @description Add colours to a WUI raster layer (No_WUI, Interface, Intermix)
#' @param .veg_perc raster layer indicating if the percentage of vegetation in
add_colours_wui <- function(.x) {

  r <- .x
  # Identify levels (1 = Interface, 2 = Intermix)
  df_lev <- data.frame(id = 0:2, wui_class = c("No_wui", "Interface", "Intermix"))
  df_lev_lst <- replicate(nlyr(r), df_lev, simplify = FALSE)
  levels(r) <- df_lev_lst
  # Add colour table 
  df_col <- data.frame(id = 0:2, col = c("lightgrey", "#FF6633", "#003366"))
  df_col_lst <- replicate(nlyr(r), df_col, simplify = FALSE)
  coltab(r) <- df_col_lst

  return(r)

}

```

## Data preprocessing

### Parameter

```{r}
#| label: Parameters-definition

raster_resolution <- 250 # m
radio_buffer <- 500 # m
area_buffer <- pi * (radio_buffer/1000)^2 # km2

```

At national scale, we have performed a raster analysis by grids cells of `r raster_resolution` m x `r raster_resolution` m.

### Study area

We load the last release of the county data from GISCO (`year = 2021`), with a resolution of `1:1 million` and `crs= 3035` (@fig-region-raster). Due to size of the file, we perform the analysis by tiles.

```{r}
#| label: fig-region-raster
#| fig-cap: Study area (Norway and Sweden)

# Create local directory for caching big files
dir.create("C:/GISCO_spatial_data")
options(gisco_cache_dir = "C:/GISCO_spatial_data")

# Counties ----
country_list <- c("NO", "SE")
counties <- giscoR::gisco_get_nuts(country = country_list,
                                   year = "2021",
                                   nuts_level = 3,
                                   epsg = "3035",
                                   resolution = "01") |> 
  filter(NUTS_NAME != "Svalbard") |> 
  filter(NUTS_NAME != "Jan Mayen") |> 
  select(NUTS_ID, NUTS_NAME, CNTR_CODE)

# Study region (SF object) ----
region_sf <- counties |> 
  summarize()

country_sf <- counties |>
  group_by(CNTR_CODE) |> 
  summarize()

# # Rogaland ----
# rogaland <- counties |> 
#   filter(NUTS_NAME == "Rogaland")

# Transform to terra::SpatVector (region_v) and SpatRaster (region_r)) -----
region_v <- vect(region_sf)
geometry_raster <- rast(region_v, resolution = raster_resolution)
region_r <- rasterize(region_v, geometry_raster)

no_v <- country_sf |> 
  filter(CNTR_CODE == "NO") |> 
  vect()

se_v <- country_sf |> 
  filter(CNTR_CODE == "SE") |> 
  vect()

# Tiles 
raster_tiles <- rast(nrows = 15,
                     ncols = 15,
                     ext(region_r),
                     crs = "EPSG:3035")

dir.create("data/processed_data")
dir.create("data/processed_data/tiles")
dir.create("data/processed_data/tiles/region")

region_tiles <- makeTiles(
  x = region_r,
  y = raster_tiles,
  filename = "data/processed_data/tiles/region/region_r_.tif",
  na.rm = TRUE,
  extend = TRUE
)

region_vrt <- vrt(region_tiles)

# Plot Study area ----

# world 
world <- rworldmap::getMap(resolution = "high") |> 
  st_as_sf() |> 
  st_transform(crs = 3035)

p1 <- ggplot() +
  geom_sf(data = world) +
  geom_sf(data = country_sf, fill = c("#D55E00", "#4E84C4")) +
  coord_sf(xlim = c(2500000, 5402285),
           ylim = c(1250000, 5500000),
           expand = FALSE) +
  theme_bw() +
  theme(panel.background = element_rect(fill = "lightgrey")) +
  theme(panel.grid = element_blank(),
        axis.text  = element_blank(),
        axis.ticks = element_blank())

p2 <- ggplot() +
  geom_sf(data = region_sf, fill = "#52854C") +
  geom_sf(data = country_sf, fill = NA) +
  theme_bw()

p2 + inset_element(p1, 0.03, 0.70, 0.6, 0.95, align_to = 'full')

```

### Building density

We have downloaded the building data from OpenStreetMaps (geofabrik: [link](https://download.geofabrik.de/europe/)), in pbf format. The version of the datasets are:

+-------------------------------------------------------------------------------------+-------------------+-----------+
| **Name**                                                                            | **Last modified** | **Size**  |
+-------------------------------------------------------------------------------------+-------------------+-----------+
| [norway-latest-osm.pbf](https://download.geofabrik.de/europe/norway-latest.osm.pbf) | 2023-08-17 04:37  | 1.2G      |
+-------------------------------------------------------------------------------------+-------------------+-----------+
| [sweden-latest-osm.pbf](https://download.geofabrik.de/europe/sweden-latest.osm.pbf) | 2023-11-16 04:50  | 656M      |
+-------------------------------------------------------------------------------------+-------------------+-----------+

They are big files and we have not put them on GitHub. If you wanted to replicate the document, you would need to download the file and save it with the same path as here, or change the path in the code (NOTE: you may have small differences in the results if OSM updates the data). To calculate the building density we do the following steps:

-   Transform the `.pbf` file to `.gpkg`, extracting only `multipolygons` (run only one time)
-   Read building polygons
-   Calculate the centroid of each building, assigning a value of 1 to the building (`building = 1`)
-   Calculate the building density in a grid cell of `r raster_resolution` m x `r raster_resolution` m (it is done by tiles). It is the `sum` of all the buildings (`building = 1`) within a grid cell
-   Merge all tiles in one raster (`terra::mosaic`)
-   Smooth the data with a circular moving window of `r radio_buffer` m; i.e., buffer area (`terra::focal`)
-   The density building of a grid cell is therefore the number of buildings in the buffer area divided by the area of the buffer. Potential WUI areas are those grid cells with a building density higher than 6.17 #/km\^2

```{r}
#| label: building-density
#| cache: true
#| warning: false
#| message: false

# # Transform OSM data to gpkg (only buildings) -- RUN ONLY ONCE 
# osm_no_pbf <- "data/big_data/OSM/norway-latest.osm.pbf"
# osm_no_building_gpkg <- osmextract::oe_vectortranslate(osm_pbf,
#                                                 layer = "multipolygons")

# # Transform OSM data to gpkg (only buildings) -- RUN ONLY ONCE
# osm_se_pbf <- "data/big_data/OSM/sweden-latest.osm.pbf"
# osm_se_building_gpkg <- osmextract::oe_vectortranslate(osm_se_pbf,
#                                                 layer = "multipolygons")

#' Function for getting buildings counts from OSM
#' @description Read only buildings from OSM.gpkg file (by tiles) and calculate building density
#' @param .tile_path Tiles where building counts should be calculated 
#' @param .data_path path to folder with OSM data (.gpkg format) 
get_building_counts <- function(.tile_path,
                                .data_path = "data/big_data/OSM"){

  # read tile
  tile <- terra::rast(.tile_path)
  files <- list.files(.data_path,
                      pattern="*.gpkg",
                      full.names = TRUE)

  # Get bounding box tile
  wkt_tile <- tile |>
    terra::project("EPSG:4326") |> 
    terra::ext() |> 
    terra::vect() |>
    terra::geom(wkt = TRUE)

  # Calculate number buildings
  bd <- lapply(files, function(.x) { sf::read_sf(.x, wkt_filter = wkt_tile) }) |>  
    dplyr::bind_rows() |> 
    tidyr::drop_na(building) |> 
    dplyr::select(building) |>
    sf::st_transform(crs = 3035) |> 
    sf::st_centroid() |> 
    dplyr::mutate(building = 1) |> 
    terra::vect() |> 
    terra::rasterize(tile, fun = sum)

  return(bd)

}

# Building density ----
build_counts <- region_tiles %>%
  # Density by tiles
  map(get_building_counts) |> 
  # Merge tiles
  sprc() |> 
  mosaic(fun = "mean")
terra::writeRaster(build_counts, 
                   "data/processed_data/build_counts.tif",
                   filetype = "GTiff", 
                   overwrite = TRUE)

# Focal (Smooth values) ----
# Number of buildings in a circle
fwm <- focalMat(build_counts, d = radio_buffer, type = "circle", fillNA = TRUE)
fwm[fwm > 0] <- 1
build_focal <- focal(build_counts, w = fwm, fun = sum , na.rm = TRUE)
terra::writeRaster(build_focal, 
                   "data/processed_data/build_focal.tif",
                   filetype = "GTiff", 
                   overwrite = TRUE)

```

```{r}
#| label: fig-building-density
#| fig-cap: Number of buildings per grid cells of 250 m x 250 m

# Read data
build_counts <- terra::rast("data/processed_data/build_counts.tif")
build_focal <- terra::rast("data/processed_data/build_focal.tif")

# Plot ----
p1 <- ggplot() +
  tidyterra::geom_spatvector(data = region_v, fill = "white") +
  tidyterra::geom_spatraster(data = build_counts) +
  scale_fill_whitebox_c(name = "N. Buildings",
                        palette = "viridi",
                        trans = "log10") +
  labs(title = "Original data from OSM") +
  theme_void()

p2 <- ggplot() +
  tidyterra::geom_spatvector(data = region_v, fill = "white") +
  tidyterra::geom_spatraster(data = build_focal) +
  scale_fill_whitebox_c(name = "N. Buildings",
                        palette = "viridi",
                        trans = "log10") +
   labs(title = "Focal values (circular window)") +
  theme_void()

p1 + p2

```

### Fuel/vegetation maps

#### EFFIS Fuel Map

The EFFIS Fuel Map (@effis_european_2017) is a pan-European map that classifies the vegetation in 10 categories according to the fire behaviour models developed by Anderson, 1982 (@anderson_aids_1982) (i.e., it excludes the last 3 categorise of the model). Resolution 250 x 250 m.

We remove small polygons (lower than 5 km^2^), and calculate the distance from each grid cell to the nearest large wild vegetation. Then, we calculate the grids that are within 2.4 km (`value = 1`) or not (`value = 0`).

We also calculate the percentage of vegetation in the buffer area (`*_wv_per`). In the study region, there are only 8 EFFIS categories (@fig-effis-fuel-map-norway).

```{r}
#| label: fig-effis-fuel-map-norway
#| fig-cap: Wildland vegetation in Norway by type of vegetation (Source EFFIS Fuel Map)
#| cache: true

# # Dowload and unzip data (RUN once)
# effis_fuel_url <- "https://effis-gwis-cms.s3-eu-west-1.amazonaws.com/effis/applications/data-and-services/FuelMap_LAEA.zip"
# dir.create("data/big_data/EFFIS")
# download.file(effis_fuel_url, "data/big_data/EFFIS/FuelMap.zip")
# unzip(zipfile = "data/big_data/EFFIS/FuelMap.zip",
#       exdir = "data/big_data/EFFIS")

# Forest layer norway
effis_fuel_path <- "data/big_data/EFFIS/FuelMap2000_NFFL_LAEA.tif"
effis_fuel <- rast(effis_fuel_path) |> 
  crop(region_v) |> 
  mask(region_v)

# Wildland vegetation
effis_wild_veg <- effis_fuel
effis_wild_veg[effis_wild_veg >= 1] <- 1
names(effis_wild_veg) <- "wild_veg"
terra::writeRaster(effis_wild_veg, 
                   "data/processed_data/effis_wild_veg.tif",
                   filetype = "GTiff", 
                   overwrite = TRUE)

# Large forest areas (> 5km2)
effis_wild_veg_large <- stars::read_stars("data/processed_data/effis_wild_veg.tif") |> 
  get_large_polygons() 
terra::writeRaster(effis_wild_veg_large, 
                   "data/processed_data/effis_wild_veg_large.tif",
                   filetype = "GTiff", 
                   overwrite = TRUE)

# Distances to large vegetated areas ----
effis_wild_veg_large_dist <- terra::distance(effis_wild_veg_large)
terra::writeRaster(effis_wild_veg_large_dist, 
                   "data/processed_data/effis_wild_veg_large_dist.tif",
                   filetype = "GTiff", 
                   overwrite = TRUE)

# Percentage-vegetation 
effis_wv_perc <- get_perc_veg_buffer(
  .veg = effis_wild_veg,
  .radio_buffer = radio_buffer)
terra::writeRaster(effis_wv_perc, 
                   "data/processed_data/effis_wv_perc.tif",
                   filetype = "GTiff", 
                   overwrite = TRUE)

# Plot ----
effis_wild_veg_list_no <- tribble(~code, ~cover_class, ~RGB,
                               1, "Short grass",	"16-240-96",
                               3, "Tall grass",	"16-128-16",
                               5, "Brush",	"255-255-0",
                               6, "Dormant brush", "255-204-0",
                               7, "Southern rough",	"255-102-0",
                               8, "Closed timber litter", "192-128-64",
                               9, "Hardwood litter", "160-64-0",
                               10, "Timber", "112-48-48") |>
  separate(RGB, c("red", "green", "blues"), sep = "-") |>  
  mutate(rgb_code = rgb(red, green, blues, maxColorValue = 255))

# Plot
effis_fuel_cls <- effis_fuel
cls <- data.frame(id = effis_wild_veg_list_no$code, 
                  value = effis_wild_veg_list_no$cover_class)
levels(effis_fuel_cls) <- cls

plot(region_r, col = "lightgrey", axes = FALSE, legend = FALSE)
plot(effis_fuel_cls, 
     col = effis_wild_veg_list_no$rgb_code,
     axes = FALSE, add = T)

```

```{r}
#| label: fig-effis-wildland-vegetation-norway
#| fig-cap: a) Wildland vegetation in Norway, and b) grid cells within a distance to large vegetation areas (>5km2) of 2.4 km (Source EFFIS Fuel Map)

# Read Effis data
effis_wild_veg <- terra::rast("data/processed_data/effis_wild_veg.tif")

# Get cells within 2400 m (YES --> 1; NO --> 0)
dist_path <- "data/processed_data/effis_wild_veg_large_dist.tif"
effis_wild_veg_large_dist_2400 <- terra::rast(dist_path) |> 
  get_dist_max(2400)

# Plot
par(mfrow = c(1,2), adj = 0)

# Wildand vegetation (Yes = 1, No = NA)
plot(region_r, col = "lightgrey", axes = F, legend = F)
plot(effis_wild_veg, col = "#009E73", axes = F, legend = F, add = T)
title("A)", adj = 0)
# Within a distance of 2.4 km (Yes = 1, No = 0)
plot(region_r, col = "lightgrey", axes = F, legend = F)
plot(effis_wild_veg_large_dist_2400, col = c(NA, "#0072B2"), axes = F, legend = F, add = T)
title("B)", adj = 0)

```

#### Corine Land Cover

We have used the Corine Land Cover 2018 (Version 2020_20u1), downloaded from [Copernicus](https://land.copernicus.eu/pan-european/corine-land-cover). The resolution of the raster is 100 m x 100 m. We classify as potential wildland vegetation the following classes @fig-corine-wild-vegetation:

-   311: [Broad-leaved forest](https://land.copernicus.eu/user-corner/technical-library/corine-land-cover-nomenclature-guidelines/html/index-clc-311.html)
-   312: [Coniferous forest](https://land.copernicus.eu/user-corner/technical-library/corine-land-cover-nomenclature-guidelines/html/index-clc-312.html)
-   313: [Mixed forest](https://land.copernicus.eu/user-corner/technical-library/corine-land-cover-nomenclature-guidelines/html/index-clc-313.html)
-   321: [Natural grassland](https://land.copernicus.eu/user-corner/technical-library/corine-land-cover-nomenclature-guidelines/html/index-clc-321.html)
-   322: [Moors and heathland](https://land.copernicus.eu/user-corner/technical-library/corine-land-cover-nomenclature-guidelines/html/index-clc-322.html)
-   323: [Sclerophyllous vegetation](https://land.copernicus.eu/user-corner/technical-library/corine-land-cover-nomenclature-guidelines/html/index-clc-323.html)
-   324: [Transitional woodland-shrub](https://land.copernicus.eu/user-corner/technical-library/corine-land-cover-nomenclature-guidelines/html/index-clc-324.html)

```{r}
#| label: fig-corine-fuel-map-norway
#| fig-cap: Wildland vegetation in Norway by type of vegetation (Source CORINE)
#| cache: true

# Load data ----
clc2018_path <- "data/big_data/corine/DATA/U2018_CLC2018_V2020_20u1.tif"

clc2018_r <- rast(clc2018_path) |> 
  crop(region_v) |> 
  mask(region_v)

# Get Wildland vegetation ----
clc2018_wild_veg_list <- tribble(~CODE_18, ~LABEL3, ~RGB,
                                 311,   "Broad-leaved forest",  "128-255-000",
                                 312,   "Coniferous forest",    "000-166-000",
                                 313,   "Mixed forest", "077-255-000",
                                 321,   "Natural grasslands",   "204-242-077",
                                 322,   "Moors and heathland",  "166-255-128",
                                 323,   "Sclerophyllous vegetation", "166-230-077",
                                 324,   "Transitional woodland-shrub",  "166-242-000") |>
  separate(RGB, c("red", "green", "blues"), sep = "-") |>  
  mutate(rgb_code = rgb(red, green, blues, maxColorValue = 255))

# Subset target vegetation
corine_fuel_veg <- clc2018_r |> 
  terra::subst(clc2018_wild_veg_list$LABEL3, clc2018_wild_veg_list$LABEL3, others = NA) |> 
  droplevels()

# Wildland vegetation: yes --> 1 
corine_wild_veg <- as.int(corine_fuel_veg)
corine_wild_veg[corine_wild_veg > 0] <- 1
# Remove colours associated with the rasters
coltab(corine_wild_veg) <- NULL
names(corine_wild_veg) <- "wild_veg"
terra::writeRaster(corine_wild_veg, 
                   "data/processed_data/corine_wild_veg.tif",
                   filetype = "GTiff", 
                   overwrite = TRUE)

# Large forest areas (> 5km2)
corine_wild_veg_large <- stars::read_stars("data/processed_data/corine_wild_veg.tif") |> 
  get_large_polygons() 
terra::writeRaster(corine_wild_veg_large, 
                   "data/processed_data/corine_wild_veg_large.tif",
                   filetype = "GTiff", 
                   overwrite = TRUE)

# Distances to large vegetated areas ----
set.values(corine_wild_veg_large)
corine_wild_veg_large_dist <- terra::distance(corine_wild_veg_large)
terra::writeRaster(corine_wild_veg_large_dist, 
                   "data/processed_data/corine_wild_veg_large_dist.tif",
                   filetype = "GTiff", 
                   overwrite = TRUE)

# Percentage-vegetation ----
corine_wv_perc <- get_perc_veg_buffer(
  .veg = corine_wild_veg,
  .radio_buffer = radio_buffer)
terra::writeRaster(corine_wv_perc, 
                   "data/processed_data/corine_wv_perc.tif",
                   filetype = "GTiff", 
                   overwrite = TRUE)

# Plot ----
plot(region_r, col = "lightgrey", axes = FALSE, legend = FALSE)
plot(corine_fuel_veg, col = clc2018_wild_veg_list$rgb_code, 
     axes = FALSE, add = T)

```

```{r}
#| label: fig-wild-vegetation-corine
#| fig-cap: a) Wildland vegetation in Norway, and b) grid cells within a distance to large vegetation areas (>5km2) of 2.4 km (Source CORINE)

# Read CORINE data
corine_wild_veg <- terra::rast("data/processed_data/corine_wild_veg.tif")

# Get cells within 2400 m (YES --> 1; NO --> 0)
dist_path <- "data/processed_data/corine_wild_veg_large_dist.tif"
corine_wild_veg_large_dist_2400 <- terra::rast(dist_path) |> 
  get_dist_max(2400)

# Plot
par(mfrow = c(1,2), adj = 0)

# Wildland vegetation (Yes = 1, No = NA)
plot(region_r, col = "lightgrey", axes = F, legend = F)
plot(corine_wild_veg, col = "#009E73", axes = F, legend = F, add = T)
title("A)", adj = 0)
# Within a distance of 2.4 km (Yes = 1, No = 0)
plot(region_r, col = "lightgrey", axes = F, legend = F)
plot(corine_wild_veg_large_dist_2400, col = c(NA, "#0072B2"), axes = F, legend = F, add = T)
title("B)", adj = 0)

```

#### National datasets

##### Norwegian Forest resource map - SR16

We downloaded the Norwegian Forest Resource map (Skogressurskart: SR16 - raster format: resolution 16 m) from [NIBIO](https://www.nibio.no/tema/skog/kart-over-skogressurser/skogressurskart-sr16). We downloaded the raster data by county (*Fylke*) in EUREF89 UTM 33, 2d (25833). There are multiple raster files but we only used the dominant tree species (*SRTRESLAG*), which is classified in three categories:

\- *gran* - spruce: 1

\- *furu* - pine: 2

\- *lauv* - deciduous: 3

We consider all of them as wildland vegetation (we reclassified the raster as 1 = wildland vegetation, and 0 = No wildland vegetation).

We aggregate the data by a factor of 6.25, so we past from an original resolution of 16x16 meters to 100x100 meters.

```{r}
#| label: SR16-fuel-map-norway
#| cache: true

# Vegetation layer norway ----
sr16r_path <- list.files("data/big_data/sr16/srtreslag",
                         pattern = "*.tif",
                         full.names = TRUE)
sr16_r <- map(sr16r_path, rast)

# Wildland vegetation
get_wild_veg <- function(.x){
    r <- .x
    r[r >= 1] <- 1 
    return(r)
  }
sr16_wild_veg <-  map(sr16_r, get_wild_veg)

# Aggregate 
change_resolution <- function(.x, .fact = 6.25) {
  aggregate(.x, fact = .fact, fun = "modal") 
}

sr16_wild_veg <- map(sr16_wild_veg, change_resolution)

# Merge rasters
sr16_wild_veg <- sr16_wild_veg |> 
  sprc() |> 
  mosaic(fun = "max") |> 
  project("EPSG:3035", method = "near")

# Export as .tif
set.values(sr16_wild_veg)
names(sr16_wild_veg) <- "wild_veg"
terra::writeRaster(sr16_wild_veg,
                   "data/processed_data/sr16_wild_veg.tif", 
                   filetype = "GTiff", 
                   overwrite = TRUE)

# Large forest areas (> 5km2)
sr16_wild_veg_large <- stars::read_stars("data/processed_data/sr16_wild_veg.tif") |>   get_large_polygons() 
set.values(sr16_wild_veg_large)
terra::writeRaster(sr16_wild_veg_large, 
                   "data/processed_data/sr16_wild_veg_large.tif",
                   filetype = "GTiff", 
                   overwrite = TRUE)

# Distances to large vegetated areas ----
sr16_wild_veg_large_dist <- terra::distance(sr16_wild_veg_large)
set.values(sr16_wild_veg_large_dist)
terra::writeRaster(sr16_wild_veg_large_dist, 
                   "data/processed_data/sr16_wild_veg_large_dist.tif",
                   filetype = "GTiff", 
                   overwrite = TRUE)

# Percentage-vegetation ----
sr16_wv_perc <- get_perc_veg_buffer(
  .veg = sr16_wild_veg,
  .radio_buffer = radio_buffer)
terra::writeRaster(sr16_wv_perc, 
                   "data/processed_data/sr16_wv_perc.tif",
                   filetype = "GTiff", 
                   overwrite = TRUE)

```

```{r}
#| label: fig-SR16-fuel-map-norway
#| fig-cap: Wildland vegetation in Norway by type of vegetation (Source SR16)

# Plot original data ----
categorical_levels <- function(.x){
  r <- .x
  levels(r) <- data.frame(id = 1:3, type =  c("spruce", "pine", "deciduous"))
  return(r)
}

sr16_vrt <- terra::vrt(sr16r_path) 
levels(sr16_vrt) <- data.frame(id = 1:3, type =  c("spruce", "pine", "deciduous"))
no_v_p <- project(no_v, "EPSG:25833")
plot(no_v_p, col = "lightgrey", axes = FALSE, legend = FALSE)
plot(sr16_vrt, col = c("lightgreen", "darkgreen", "yellow"), axes = FALSE, add = T)

```

##### Swedish National Land Cover Database (NMD)

We download the NMD Base map from the Swedish Environmental Protection Agency ([Nationella marktäckedata 2018; basskikt](https://geodata.naturvardsverket.se/nedladdning/marktacke/NMD2018/NMD2018_basskikt_ogeneraliserad_Sverige_v1_1.zip)). It classified Sweden in 6 main levels, which later are subdivided in level 2 and level 3 (@fig-NMD-veg-map-sewden):

+----------------------+----------------------------------------------------+--------------------------------------------------------+
| Level 1              | Level 2                                            | Level 3                                                |
+======================+====================================================+========================================================+
| 1.  Skog             | 1.1 Skog utanför våtmark                           | 1.1.1 Tallskog utanför våtmark                         |
|                      |                                                    |                                                        |
|                      |                                                    | 1.1.2 Granskog utanför våtmark                         |
|                      |                                                    |                                                        |
|                      |                                                    | 1.1.3 Barrblandskog utanför våtmark                    |
|                      |                                                    |                                                        |
|                      |                                                    | 1.1.4 Lövblandad barrskog utanför våtmark              |
|                      |                                                    |                                                        |
|                      |                                                    | 1.1.5 Triviallövskog utanför våtmark                   |
|                      |                                                    |                                                        |
|                      |                                                    | 1.1.6 Ädellövskog utanför våtmark                      |
|                      |                                                    |                                                        |
|                      |                                                    | 1.1.7 Triviallövskog med ädellövinslag utanför våtmark |
|                      |                                                    |                                                        |
|                      |                                                    | 1.1.8 Temporärt ej skog utanför våtmark                |
+----------------------+----------------------------------------------------+--------------------------------------------------------+
|                      | 1.2 Skog på våtmark                                | 1.2.1 Tallskog på våtmark                              |
|                      |                                                    |                                                        |
|                      |                                                    | 1.2.2 Granskog på våtmark                              |
|                      |                                                    |                                                        |
|                      |                                                    | 1.2.3 Barrblandskog på våtmark                         |
|                      |                                                    |                                                        |
|                      |                                                    | 1.2.4 Lövblandad barrskog på våtmark                   |
|                      |                                                    |                                                        |
|                      |                                                    | 1.2.5 Triviallövskog på våtmark                        |
|                      |                                                    |                                                        |
|                      |                                                    | 1.2.6 Ädellövskog på våtmark                           |
|                      |                                                    |                                                        |
|                      |                                                    | 1.2.7 Triviallövskog med ädellövinslag på våtmark      |
|                      |                                                    |                                                        |
|                      |                                                    | 1.2.8 Temporärt ej skog på våtmark                     |
+----------------------+----------------------------------------------------+--------------------------------------------------------+
| 2.  Våtmark          |                                                    |                                                        |
+----------------------+----------------------------------------------------+--------------------------------------------------------+
| 3.  Åkermark         |                                                    |                                                        |
+----------------------+----------------------------------------------------+--------------------------------------------------------+
| 4.  Övrig öppen mark | 4.1 Övrig öppen mark utan vegetation               |                                                        |
+----------------------+----------------------------------------------------+--------------------------------------------------------+
|                      | 4.2 Övrig öppen mark med vegetation                |                                                        |
+----------------------+----------------------------------------------------+--------------------------------------------------------+
| 5.  Exploaterad mark | 5.1 Exploaterad mark, byggnad                      |                                                        |
+----------------------+----------------------------------------------------+--------------------------------------------------------+
|                      | 5.2 Exploaterad mark, ej byggnad eller väg/järnväg |                                                        |
+----------------------+----------------------------------------------------+--------------------------------------------------------+
|                      | 5.3 Exploaterad mark, väg/järnväg                  |                                                        |
+----------------------+----------------------------------------------------+--------------------------------------------------------+
| 6.  Vatten           | 6.1 Sjö och vattendrag                             |                                                        |
+----------------------+----------------------------------------------------+--------------------------------------------------------+
|                      | 6.2 Hav                                            |                                                        |
+----------------------+----------------------------------------------------+--------------------------------------------------------+

: NMD categories (table Bilaga 1: Nomenklatur)

```{r}
#| label: fig-NMD-veg-map-sewden
#| fig-cap: Vegetation map of Sweden (National Land Cover Database 2018)

nmd_path <- "data/big_data/NMD2018/nmd2018bas_ogeneraliserad_v1_1.tif" 
nmd_r <- rast(nmd_path)

# Vegetation code
levles_lst <- levels(nmd_r)
levles_lst[[1]]$Klass <- iconv(levles_lst[[1]]$Klass, "latin1")
levels(nmd_r) <- levles_lst

nmd_veg_list <- levles_lst[[1]] |>
  as_tibble() |> 
  mutate(Klass = ifelse(Klass == "", NA, Klass)) |> 
  drop_na()

plot(nmd_r)

```

Then, we classified as Non-burnable the following categories:

```{r}
#| label: tbl-nmd-burnable
#| tbl-cap: Non-burnable vegetation (categories)

nmd_no_fuel_list <- nmd_veg_list |> 
  filter(value %in% c(3, 41, 42, 51, 52, 53, 61, 62, 131, 169, 170, 179, 181, 189, 190))

nmd_no_fuel_list |> 
  gt()

```

The original resolution of the map is 10 meters. We aggregate it (by a factor of 10) and transform the coordinate system from `SWEREF99 TM (EPSG:3006)` to `ETRS89-extended / LAEA Europe (EPSG:3035)`.

```{r}
#| label: NMD-wild-veg-map-sewden
#| cache: true

# Reduce resolution 
nmd_r_a <- aggregate(nmd_r, fact = 10, fun = "modal") |> 
  project("EPSG:3035", method = "near")

# Remove no fuel vegetations 
nmd_wild_veg <- nmd_r_a |> 
  mutate(Klass = ifelse(Klass %in% nmd_no_fuel_list$Klass, NA, Klass))

# Wild vegetation; YES = 1, NO = 0
nmd_wild_veg[nmd_wild_veg == 1] <- NA  # Outside Sweden 
nmd_wild_veg[nmd_wild_veg  > 1] <- 1

# Export as .tif
names(nmd_wild_veg) <- "wild_veg"
coltab(nmd_wild_veg) <- NULL
set.values(nmd_wild_veg)
terra::writeRaster(nmd_wild_veg,
                   "data/processed_data/nmd_wild_veg.tif", 
                   filetype = "GTiff", 
                   overwrite = TRUE)

# Large forest areas (> 5km2)
nmd_wild_veg_large <- stars::read_stars("data/processed_data/nmd_wild_veg.tif") |> 
  get_large_polygons() 
set.values(nmd_wild_veg_large)
terra::writeRaster(nmd_wild_veg_large, 
                   "data/processed_data/nmd_wild_veg_large.tif",
                   filetype = "GTiff", 
                   overwrite = TRUE)

# Distances to large vegetated areas ----
nmd_wild_veg_large_dist <- terra::distance(nmd_wild_veg_large)
set.values(nmd_wild_veg_large_dist)
terra::writeRaster(nmd_wild_veg_large_dist, 
                   "data/processed_data/nmd_wild_veg_large_dist.tif",
                   filetype = "GTiff", 
                   overwrite = TRUE)

# Percentage-vegetation ----
nmd_wv_perc <- get_perc_veg_buffer(
  .veg = nmd_wild_veg,
  .radio_buffer = radio_buffer)
set.values(nmd_wv_perc)
terra::writeRaster(nmd_wv_perc, 
                   "data/processed_data/nmd_wv_perc.tif",
                   filetype = "GTiff", 
                   overwrite = TRUE)

```

Finally, we combine both maps in one (@fig-wild-vegetation-national)

```{r}
#| label: fig-wild-vegetation-national
#| fig-cap: a) Wildland vegetation in the region, and b) grid cells within a distance to large vegetation areas (>5km2) of 2.4 km (Source SR16 and NMD)

# Read SR16 data
sr16_wild_veg <- terra::rast("data/processed_data/sr16_wild_veg.tif")
dist_path <- "data/processed_data/sr16_wild_veg_large_dist.tif"
sr16_wild_veg_large_dist_2400 <- terra::rast(dist_path) |> 
  get_dist_max(2400)

# Read NRM data
nmd_wild_veg <- terra::rast("data/processed_data/nmd_wild_veg.tif")
dist_path <- "data/processed_data/nmd_wild_veg_large_dist.tif"
nmd_wild_veg_large_dist_2400 <- terra::rast(dist_path) |> 
  get_dist_max(2400)

# Plot
par(mfrow = c(1,2), adj = 0)

# Wildland vegetation (Yes = 1, No = NA)
plot(region_r, col = "lightgrey", axes = F, legend = F)
plot(sr16_wild_veg, col = "#009E73", axes = F, legend = F, add = T)
plot(nmd_wild_veg, col = "#009E73", axes = F, legend = F, add = T)
title("A)", adj = 0)
# Within a distance of 2.4 km (Yes = 1, No = 0)
plot(region_r, col = "lightgrey", axes = F, legend = F)
plot(sr16_wild_veg_large_dist_2400, 
     col = c(NA, "#0072B2"),
     axes = F, 
     legend = F,
     add = T)
plot(nmd_wild_veg_large_dist_2400, 
     col = c(NA, "#0072B2"), 
     axes = F, 
     legend = F, 
     add = T)
title("B)", adj = 0)

```

#### FirEUrisk_Europe_fuel_map

Original data from [@Aragoneses2022]. Al fuels are considered for WUI. The resolution of the original map is 1km and we have re-sampled it to a grid cell of 250x250 grids (same than `region_r`) for the WUI (based in the type that is most presented in the area: `mode`).

```{r}
#| label: fig-firEUrisk-fuel-map-norway
#| fig-cap: Wildland vegetation in Norway by type of vegetation (Source EFFIS Fuel Map)
#| cache: true

# Codes
firEUrisk_fuel_type <- tribble(~code, ~description,
                                1111,"Open broadleaf evergreen forest",
                                  23, "High shrubland",
                                1112, "Closed broadleaf evergreen forest",
                                  31, "Low grassland",
                                1121, "Open broadleaf deciduous forest",
                                  32, "Medium grassland",
                                1122, "Closed broadleaf deciduous forest",
                                  33, "High grassland",
                                1211, "Open needleleaf evergreen forest",
                                  41, "Herbaceous cropland",
                                1212, "Closed needleleaf evergreen forest",
                                  42, "Woody cropland",
                                1221, "Open needleleaf deciduous forest",
                                  51, "Wet and peat/semi-peat land – tree",
                                1222, "Closed needleleaf deciduous forest", 
                                  52, "Wet and peat/semi-peat land – shrubland",
                                1301, "Open mixed forest",
                                  53, "Wet and peat/semi-peat land – grassland",
                                1302, "Closed mixed forest",
                                  61, "Urban continuous fabric",
                                  21, "Low shrubland",
                                  62, "Urban discontinuous fabric",
                                  22, "Medium shrubland",
                                   7, "Nonfuel")

firEUrisk_no_fuel_type <- tribble(~code, ~description,
                                     41, "Herbaceous cropland",
                                     42, "Woody cropland",
                                     51, "Wet and peat/semi-peat land – tree",
                                     52, "Wet and peat/semi-peat land – shrubland",
                                     53, "Wet and peat/semi-peat land – grassland",
                                     61, "Urban continuous fabric",
                                     62, "Urban discontinuous fabric",
                                      7, "Nonfuel")

# Forest layer norway
firEUrisk_fuel_path <- "data/big_data/firEUrisk/FirEUrisk_Europe_fuel_map.tif"
firEUrisk_fuel <- rast(firEUrisk_fuel_path) |> 
  crop(region_v) |> 
  mask(region_v)

# Resample to grid cells of 250m x 250m (same than region_r)
firEUrisk_fuel_250m <- resample(firEUrisk_fuel, region_r, method = "mode") |> 
 # Remove no fuel vegetations 
  mutate(FirEUrisk_Europe_fuel_map = ifelse(FirEUrisk_Europe_fuel_map %in% firEUrisk_no_fuel_type$code,
                    NA,
                    FirEUrisk_Europe_fuel_map))

# Wildland vegetation
firEUrisk_wild_veg <- firEUrisk_fuel_250m 
firEUrisk_wild_veg[firEUrisk_wild_veg >= 1] <- 1
names(firEUrisk_wild_veg) <- "wild_veg"
terra::writeRaster(firEUrisk_wild_veg, 
                   "data/processed_data/firEUrisk_wild_veg.tif",
                   filetype = "GTiff", 
                   overwrite = TRUE)

# Large forest areas (> 5km2)
firEUrisk_wild_veg_large <- stars::read_stars("data/processed_data/firEUrisk_wild_veg.tif") |> 
  get_large_polygons() 
terra::writeRaster(firEUrisk_wild_veg_large, 
                   "data/processed_data/firEUrisk_wild_veg_large.tif",
                   filetype = "GTiff", 
                   overwrite = TRUE)

# Distances to large vegetated areas ----
set.values(firEUrisk_wild_veg_large)
firEUrisk_wild_veg_large_dist <- terra::distance(firEUrisk_wild_veg_large)
terra::writeRaster(firEUrisk_wild_veg_large_dist, 
                   "data/processed_data/firEUrisk_wild_veg_large_dist.tif",
                   filetype = "GTiff", 
                   overwrite = TRUE)

# Percentage-vegetation ----
firEUrisk_wv_perc <- get_perc_veg_buffer(
  .veg = firEUrisk_wild_veg,
  .radio_buffer = radio_buffer)
terra::writeRaster(firEUrisk_wv_perc, 
                   "data/processed_data/firEUrisk_wv_perc.tif",
                   filetype = "GTiff", 
                   overwrite = TRUE)
# Plot ----
levels(firEUrisk_fuel_250m) <- data.frame(id = firEUrisk_fuel_type$code,
                                          type =  firEUrisk_fuel_type$description)

plot(region_r, col = "lightgrey", axes = FALSE, legend = FALSE)
plot(firEUrisk_fuel_250m, axes = FALSE, add = T)

```

```{r}
#| label: fig-firEUrisk-wildland-vegetation-norway
#| fig-cap: a) Wildland vegetation in Norway, and b) grid cells within a distance to large vegetation areas (>5km2) of 2.4 km (Source EFFIS Fuel Map)

# Read Effis data
firEUrisk_wild_veg <- terra::rast("data/processed_data/firEUrisk_wild_veg.tif")
dist_path <- "data/processed_data/firEUrisk_wild_veg_large_dist.tif"
firEUrisk_wild_veg_large_dist_2400 <- terra::rast(dist_path) |> 
  get_dist_max(2400)

# Plot
par(mfrow = c(1,2), adj = 0)

# Wildand vegetation (Yes = 1, No = NA)
plot(region_r, col = "lightgrey", axes = F, legend = F)
plot(firEUrisk_wild_veg, col = "#009E73", axes = F, legend = F, add = T)
title("A)", adj = 0)
# Within a distance of 2.4 km (Yes = 1, No = 0)
plot(region_r, col = "lightgrey", axes = F, legend = F)
plot(firEUrisk_wild_veg_large_dist_2400, col = c(NA, "#0072B2"), axes = F, legend = F, add = T)
title("B)", adj = 0)

```

## WUI classification

We have used the definition described by [@stewart_defining_2007; @bar-massada_comparative_2021; @bar-massada_using_2013], which is shown in [@fig-WUI-raster; @bar-massada_comparative_2021]:

```{r}
#| label: fig-WUI-raster 
#| fig-cap: "Zonal-based WUI mapping approach"

knitr::include_graphics("figures/wui_map_approach_bar_massada_2021.png")

```

We carry out a point-based approach with grid cells of `r raster_resolution` m x `r raster_resolution` m, and a buffer radio of `r radio_buffer` m. Our workflow is therefore:

```{mermaid}
flowchart TD
    A["Grid 250m x 250m"] --> B{"Building density > 6.17 #/km2"}
    B -- Yes --> C{"Vegetation > 50%"}
    B -- No --> D["No WUI"]
    C -- Yes --> E["Intermix"]
    C -- No --> F{"Within 2.4 km of large 
    forest (>5 km)"}
    F -- Yes --> G["Interface"]
    F -- No --> H["No WUI"]
    style D fill:lightgrey
    style E fill:#003366,color:#FFFFFF
    style G fill:#FF6633
    style H fill:lightgrey

```

Therefore, we do the following steps:

1.  Create a raster with the potential WUI cells (`wui_potential`)

-   Building density (in the buffer area) ≤ 6.16 #/km\^2 --\> NA

-   Building density (in the buffer area) \> 6.16 #/km\^2 --\> 1

2.  Create a raster with the grids where the percentage of fuel vegetation in the buffer area is higher than 50% (`wv_perc`)

-   Percentage \< 50% --\> 0
-   Percentage ≥ 50% --\> 1

3.  Create a raster indicating if the grid cell is within 2.4 km of a large (\>5 km2) fuel vegetation area (`wild_veg_large_dist_*`)

-   Yes: 1
-   No: 0

4.  Sum both rasters (`wv_perc + wui_potential`), and generate a new raster (`wui`) with three options :

-   wui = NA --\> No WUI

-   wui = 2 --\> WUI intermix

-   wui = 1 --\> Potential interface (it is reassigned to 0 in the raster)

5.  Sum this raster (`wui`) and `wild_veg_large_dist_`, we get therefore that:

-   wui = NA --\> No WUI
-   wui = 0 --\> No WUI
-   wui = 1 --\> Interface
-   wui = 2 --\> Intermix
-   wui = 3 --\> Intermix

```{r}
#| label: fig-potential-wui
#| fig-cap: Potential WUI grids based on building density

# Minimum number of buildings (N) in the buffer area ----
# WUI if building density (d) > 6.17 [#/km^2]
# N = d [#/km^2] * A [km^2]
# A = pi * radio^2
d_min <- 6.17
N_min <- round(d_min * area_buffer, 3)

wui_potential <- build_focal
wui_potential[wui_potential <= N_min] <- NA
wui_potential[wui_potential >  N_min] <- 1

plot(region_r, col = "lightgrey", axes = F, legend = F)
plot(wui_potential, axes = F, legend = F, add = T)

```

```{r}
#| label: read-wv-data

effis_wv_perc     <- terra::rast("data/processed_data/effis_wv_perc.tif")
corine_wv_perc    <- terra::rast("data/processed_data/corine_wv_perc.tif")
sr16_wv_perc      <- terra::rast("data/processed_data/sr16_wv_perc.tif")
nmd_wv_perc       <- terra::rast("data/processed_data/nmd_wv_perc.tif")
firEUrisk_wv_perc <- terra::rast("data/processed_data/firEUrisk_wv_perc.tif")

```

```{r}
#| label: calculate-wui-region
#| cache: true

# EFFIS
effis_wui <- get_wui_raster(.veg_perc = effis_wv_perc,
                            .veg_dist = effis_wild_veg_large_dist_2400)

# CORINE
corine_wui <- get_wui_raster(.veg_perc = corine_wv_perc,
                             .veg_dist = corine_wild_veg_large_dist_2400)

# National maps
sr16_wui <- get_wui_raster(.veg_perc = sr16_wv_perc,
                           .veg_dist = sr16_wild_veg_large_dist_2400,
                           .region = no_v)

nrm_wui <- get_wui_raster(.veg_perc = nmd_wv_perc,
                          .veg_dist = nmd_wild_veg_large_dist_2400,
                          .region = se_v)

national_wui <- mosaic(sr16_wui, nrm_wui, fun = "max") |> 
  add_colours_wui()

# firEUrisk
firEUrisk_wui <- get_wui_raster(.veg_perc = firEUrisk_wv_perc,
                                .veg_dist = firEUrisk_wild_veg_large_dist_2400)

# Put all rasters in one 
wui_r <- c(effis_wui, corine_wui, national_wui, firEUrisk_wui)
names(wui_r) <- c("effis", "corine", "national", "firEUrisk")

terra::saveRDS(wui_r, "data/processed_data/wui_r.rds")

```

Map of WUI areas (@fig-wui-maps-region; @tbl-pct-wui-region).

```{r}
#| label: read-wui-raster

wui_r <- terra::readRDS("data/processed_data/wui_r.rds")

```

```{r}
#| label: fig-wui-maps-region
#| fig-cap: "WUI maps of Norway based on different fuel/vegetation maps (resolution of 250 x 250 m)"
#| fig-subcap: 
#|   - "EFFIS"
#|   - "CORINE"
#|   - "FirEUrisk"
#|   - "National databases"
#| layout-ncol: 2
#| fig-width: 4
#| fig-format: png
#| cache: true

f_plot <- function(.x){ plot(.x, 
                             axes = F, 
                             mar = c(0, 0, 0, 0),
                             legend = "topleft")  }

f_plot(wui_r$effis)
f_plot(wui_r$corine)
f_plot(wui_r$firEUrisk)
f_plot(wui_r$national)

```

Percentages at national scale from the different vegetation/fuel types (based on the number of grids cells of `r raster_resolution`x `r raster_resolution` \[m\] for each WUI type):

```{r}
#| label: tbl-pct-wui-region
#| tbl-cap: Percentages of WUI by vegetation and country
#| tbl-subcap: 
#|  - Norway
#|  - Sweden
#| layout-ncol: 2

# summary table 
tbl_summary_vegetation <- function(.x){

  .x |> 
    as_tibble() |> 
    drop_na() |> 
    pivot_longer(everything(), 
                 names_to = "name",
                 values_to = "value") |> 
    group_by(name, value) |> 
    summarise(count = n()) |> 
    ungroup() |> 
    group_by(name) |> 
    mutate(total = sum(count),
           pct = 100*count/total) |> 
    pivot_wider(names_from = "value", 
                values_from = c("count", "pct")) |> 
    mutate(pct_wui = pct_Interface + pct_Intermix) |> 
    select(starts_with("pct")) |> 
    relocate(pct_wui, .after = pct_No_wui) |> 
    gt(row_group_as_column = TRUE) |> 
    fmt_number(decimals = 1) |> 
    cols_label(
      name = "",
      pct_No_wui = md("**No WUI**"),
      pct_Interface = md("**Interface**"),
      pct_Intermix = md("**Intermix**"),
      pct_wui = md("**Total**")
    )  |> 
    tab_spanner(
      label = md("**WUI**"),
      columns = c(pct_Interface, pct_Intermix, pct_wui)
    ) |> 
    tab_options(
      table.font.size = 9,
      row_group.padding = px(1),
      row_group.font.weight = "bold",
      data_row.padding = px(1)
    )

}

# Summary WUI by country 
wui_r_no <- mask(wui_r, no_v)
wui_r_se <- mask(wui_r, se_v)

# Tables
tbl_summary_vegetation(wui_r_no) 
tbl_summary_vegetation(wui_r_se)

```

Zoom to counties (@fig-wui-map-counties; @tbl-summary-wui-county):

```{r}
#| label: fig-wui-map-counties
#| fig-cap: "WUI maps by county based on different fuel/vegetation maps (resolution of 250 x 250 m)"
#| fig-subcap: 
#|  - "Viken"
#|  - "Rogaland"
#|  - "Vestland"
#|  - "Trøndelag"
#|  - "Troms og Finnmark"
#|  - "Innlandet"
#|  - "Oslo"
#|  - "Vestfold og Telemark"
#|  - "Agder"
#|  - "Møre og Romsdal"
#|  - "Nordland"
#| layout-nrow: 11
#| out-width: 100%

# plot_zoom <- function(.x) {
#   
#   zoom_area <- counties |> 
#     filter(NUTS_NAME == {{ .x }})
#   
#   wui_r_zoom <- wui_r |> 
#     crop(zoom_area) |> 
#     mask(zoom_area)
#   
#   # Add colour table 
#   coltab(wui_r_zoom) <- 
#     data.frame(id = 0:2, 
#                col = c("lightgrey", "#F0E442", "#CC79A7"))
#   
#   par(mfrow = c(1, 4))
#   
#   plot(wui_r_zoom$effis, 
#        axes = FALSE, 
#        font.main = 1,
#        legend = FALSE,
#        mar = c(0, 0, 0, 0),
#        main =  names(wui_r_zoom$effis))
#   # title( {{ .x }}, line = -4, adj = 0)
#   
#   plot(wui_r_zoom$corine, 
#        axes = FALSE, 
#        font.main = 1,
#        legend = FALSE,
#        mar = c(0, 0, 0, 0),
#        main =  names(wui_r_zoom$corine))
#   
#   plot(wui_r_zoom$firEUrisk,
#        axes = FALSE, 
#        font.main = 1,
#        legend = FALSE,
#        mar = c(0, 0, 0, 0), 
#        main =  names(wui_r_zoom$firEUrisk))
#   
#   plot(wui_r_zoom$sr16,
#        axes = F, 
#        font.main = 1,
#        legend = FALSE,
#        mar = c(0, 0, 0, 0), 
#        main =  names(wui_r_zoom$sr16)) 
#   
# }
# 
# # par(mfrow = c(1, 11))
# as.list(counties$NUTS_NAME) |> 
#   walk(plot_zoom)

```

```{r}
#| label: tbl-summary-wui-county
#| tbl-cap: Percentages of WUI by county
#| tbl-subcap: 
#|  - Norway
#|  - Sweden
#| layout-ncol: 2

# Function for printing the table
tbl_summary_county <- function(.x) {
  .x |> 
    select(NUTS_NAME, name, starts_with("pct")) |> 
    gt(rowname_col = "name",
       groupname_col = "NUTS_NAME",
       row_group_as_column = FALSE) |> 
    fmt_number(decimals = 1) |> 
    cols_label(
      name = "",
      pct_No_wui = md("**No WUI**"),
      pct_Interface = md("**Interface**"),
      pct_Intermix = md("**Intermix**"),
      pct_wui = md("**Total**")
    )  |> 
    tab_spanner(
      label = md("**WUI**"),
      columns = c(pct_Interface, pct_Intermix, pct_wui)
    ) |> 
    tab_options(
      table.font.size = 9,
      row_group.background.color = "lightgrey",
      row_group.padding = px(1),
      row_group.font.weight = "bold",
      data_row.padding = px(1)
    ) 
}

# Summary by county
tbl_wui_county <- terra::extract(wui_r, counties) |>  
  pivot_longer(cols = -ID, 
               names_to = "name",
               values_to = "value") |> 
  group_by(ID, name, value) |> 
  summarise(count = n()) |> 
  ungroup() |> 
  group_by(ID, name) |> 
  mutate(total = sum(count), 
         pct = 100*count/total) |> 
  ungroup() |> 
  pivot_wider(names_from = "value",
              values_from = c("count", "pct")) |> 
  mutate(pct_wui = pct_Interface + pct_Intermix)

# Add values to sf counties
counties_wui <- counties |> 
  st_as_sf()  %>% 
  mutate(ID := seq_len(nrow(.)))  |> 
  left_join(tbl_wui_county, by = "ID") |> 
  st_drop_geometry() |> 
  as_tibble()

# Table by country & county 
filter(counties_wui, CNTR_CODE == "NO") |> 
  tbl_summary_county()

filter(counties_wui, CNTR_CODE == "SE") |> 
  tbl_summary_county()

```
```{r}
#| label: fig-pct-wui-counties

# Add county ID
counties_cent_ID <- counties |>
  st_centroid() %>%
  mutate(x = st_coordinates(.)[,1],
         y = st_coordinates(.)[,2]) |> 
  arrange(desc(y), x) |> 
  mutate(ID = seq_along(counties$NUTS_NAME))

counties_ID <- counties |> 
  as_tibble() |> 
  left_join(counties_cent_ID, "NUTS_NAME") |> 
  st_sf()

# geofacet

mygrid <- data.frame(
  code = c("1 Troms og F.", 
           "2 Norrbotten", 
           "3 Nordland", 
           "4 Västerbotten", 
           "5 Trøndelag", 
           "7 Västernorrland", 
           "6 Jämtland", 
           "8 Møre og R.", 
           "10 Innlandet", 
           "11 Dalarna", 
           "9 Gävleborg", 
           "12 Vestland",
           "15 Oslo",
           "14 Viken",
           "13 Uppsala",
           "16 Västmanland",
           "20 Örebro", 
           "17 Värmland",
           "22 Rogaland", 
           "19 Vestfold og T.",
           "18 Stockholm", 
           "21 Södermanland",
           "24 Östergötland",
           "25 Västra Götaland",
           "23 Agder", 
           "29 Halland", 
           "27 Jönköping", 
           "28 Kalmar", 
           "26 Gotland", 
           "30 Kronoberg", 
           "31 Blekinge",
           "32 Skåne"),
  row = c(2, 2, 3, 3, 4, 4, 4, 5, 5, 5,
          5, 6, 6, 6, 6, 6, 6, 6, 7, 7, 
          7, 7, 7, 7, 8, 8, 8, 8, 9, 9, 9, 10),
  col = c(6, 7, 6, 7, 5, 7, 6, 4, 5, 6,
          7, 3, 5, 4, 9, 8, 7, 6, 3, 4, 
          9, 8, 7, 6, 4, 6, 7, 8, 9, 6, 7, 6),
  name = c("Troms og Finnmark",
           "Norrbotten", 
           "Nordland",
           "Västerbotten", 
           "Trøndelag", 
           "Västernorrland",
           "Jämtland", 
           "Møre og Romsdal",
           "Innlandet", 
           "Dalarna", 
           "Gävleborg",
           "Vestland", 
           "Oslo", 
           "Viken", 
           "Uppsala",
           "Västmanland",
           "Örebro", 
           "Värmland",
           "Rogaland", 
           "Vestfold og Telemark",
           "Stockholm", 
           "Södermanland", 
           "Östergötland", 
           "Västra Götaland", 
           "Agder", 
           "Halland", 
           "Jönköping", 
           "Kalmar",
           "Gotland", 
           "Kronoberg", 
           "Blekinge", 
           "Skåne"),
  stringsAsFactors = FALSE
)

# Maps

p_map <- ggplot() +
  geom_sf(data = counties_ID, 
          fill = "#FAFAFA",
          col = "lightgrey",
          size = 0.5) +
  geom_sf_text(data = counties_ID, 
               aes(label = ID),
               col = "red",
               size = 3) +
  theme_void()

p_bar <- counties_wui |> 
  select(NUTS_NAME, name, starts_with("pct")) |> 
  select(-pct_wui) |> 
  pivot_longer(cols = starts_with("pct"), 
               names_to = "type",
               values_to = "value") |> 
  mutate(type = str_remove_all(type, "pct_"),
         NUTS_NAME = str_remove_all(NUTS_NAME, "s län"),
         NUTS_NAME = str_remove_all(NUTS_NAME, " län")) |> 
  mutate(name = case_when(name == "corine" ~ "C",
                          name == "effis" ~ "E",
                          name == "national" ~ "N",
                          name == "firEUrisk" ~ "F")) |>
  mutate(name = factor(name, levels = c("N", "C", "E", "F"))) |> 
  mutate(type = factor(type, levels = c("No_wui", "Interface", "Intermix"))) |> 
  ggplot() +
  geom_bar(aes(x = name,
               y = value,
               fill = type),
           col = "grey",
           stat = "identity",
           position = "fill"
           ) +
  scale_fill_manual(values = c("#FAFAFA", "#FF6633", "#003366")) +
  geofacet::facet_geo(~NUTS_NAME,
                      grid = mygrid, 
                      label = "code"
                      ) +
  labs(x = "",
       y = "") +
  theme_bw() +
  theme(
    strip.text = element_text(size = 9, face = "bold")
  )

ggsave(plot = p_map, 
       filename = "figures/counties_map.png", 
       width = 3, height = 3)

ggsave(plot = p_bar, 
       filename = "figures/counties_wui_perc.png", 
       width = 15, height = 20)

```

![](figures/counties_map.png)
![WUI percentage by county](figures/counties_wui_perc.png)

Area of wild vegetation by county (@tbl-summary-table-areas): Number vegetation cells (centroid is inside the polygon), sum then and calculate the area based on the raster resolution.

```{r}
#| label: tbl-summary-table-areas
#| tbl-cap: Number of buildings and wildland vegetation percentage by county (and total)
#| cache: true

# Number of buildings by county  ----
county_build_count <- build_counts |> 
  # Number of buildings per county
  extract(y = vect(counties),
          fun = "sum",
          na.rm = TRUE,
          bind = TRUE) |> 
  # sf object
  st_as_sf() %>% 
  # Area polygons
  mutate(area = units::set_units(st_area(.), "km2"))  |> 
  # Building density by county
  mutate(build_den = sum/area)

# Percentage vegetation by county ----

# Function for getting vegetation area by counties 
f_pct_veg <- function(df) {
  df |> # sf object with vegetation data  
    # Number of cells
    extract(y = vect(counties),
            fun = "sum",
            na.rm = TRUE,
            bind = TRUE) |>  
    st_as_sf() |> 
    # Area (number cells x area cells)
    mutate(
      area_km = wild_veg * ((res(df))[1]/1000 * (res(df ))[2]/1000)
    ) |> 
    # Get only values 
    st_drop_geometry() |> 
    select(NUTS_NAME, CNTR_CODE, area_km)
}

## Run for each vegetation type

# Effis
effis_county  <- f_pct_veg(effis_wild_veg) |> 
    rename(area_km_effis = area_km)

# Corine
corine_county <- f_pct_veg(corine_wild_veg) |> 
  rename(area_km_corine = area_km)

# National
sr16_county <- f_pct_veg(sr16_wild_veg) |> 
  filter(CNTR_CODE  == "NO")
nmd_county  <- f_pct_veg(nmd_wild_veg)  |> 
  filter(CNTR_CODE  == "SE")

national_county <- rbind(sr16_county, nmd_county) |> 
  rename(area_km_national = area_km)

# FirEUrisk
firEUrisk_county <- f_pct_veg(firEUrisk_wild_veg) |> 
  rename(area_km_firEUrisk = area_km)

# Joint tables
county_veg <- national_county |> 
  left_join(corine_county) |> 
  left_join(effis_county) |> 
  left_join(firEUrisk_county)

# Add vegetation to the summary table ----
df_summary <- county_build_count |> 
  st_drop_geometry() |> 
  left_join(county_veg) |> 
  mutate(across(.cols = starts_with("area_km"),
                .fns = ~ 100 * (.x / as.numeric(area)),
                .names = "perc_{.col}"))

# Summary tables by country and county ----
summary_country <- df_summary |> 
  select(CNTR_CODE, sum, area, starts_with("area_km")) |> 
  group_by(CNTR_CODE) |> 
  summarise_all(sum) |> 
  mutate(NUTS_NAME = "Total",
         build_den = sum/area) |> 
  mutate(across(.cols = starts_with("area_km"),
                .fns = ~ 100 * (.x / as.numeric(area)),
                .names = "perc_{.col}")) |> 
  relocate(build_den, .after = area) |> 
  relocate(NUTS_NAME, .after = CNTR_CODE) |> 
  select(CNTR_CODE, NUTS_NAME, area,
         sum, build_den,
         starts_with("perc")) |> 
  mutate(area = as.numeric(area),
         build_den = as.numeric(build_den),
         CNTR_CODE = case_when(CNTR_CODE == "NO" ~ "NORWAY",
                               CNTR_CODE == "SE" ~ "SWEDEN"))

summary_county <- df_summary |> 
  select(CNTR_CODE, NUTS_NAME, area,
         sum, build_den,
         starts_with("perc")) |>
  arrange(CNTR_CODE, NUTS_NAME) |> 
  mutate(area = as.numeric(area),
         build_den = as.numeric(build_den),
         CNTR_CODE = case_when(CNTR_CODE == "NO" ~ "NORWAY",
                               CNTR_CODE == "SE" ~ "SWEDEN"))

# Print table ----
summary_county |> 
  rbind(summary_country) |> 
  gt(groupname_col = "CNTR_CODE",
     row_group_as_column = FALSE) |> 
  fmt_number(c(build_den, starts_with("perc")), decimals = 1) |> 
  fmt_number(c(sum, area), decimals = 0) |> 
  tab_options(
      table.font.size = 9,
      row_group.background.color = "lightgrey",
      row_group.padding = px(1),
      row_group.font.weight = "bold",
      data_row.padding = px(1)
    ) |> 
  tab_spanner(
      label = md("**Area covered by vegetation [%]**"),
      columns = starts_with("perc")
    ) |> 
  tab_spanner(
      label = md("**Buildings**"),
      columns = c(sum, build_den)
    ) |> 
  cols_label(
      NUTS_NAME = md("**County**"),
      area = md("**Area [km2]**"),
      sum = md("**Count**"),
      build_den = md("**Density [#/km2]**"),
      perc_area_km_national = md("**National**"),
      perc_area_km_effis = md("**Effis**"),
      perc_area_km_corine = md("**Corine**"),
      perc_area_km_firEUrisk = md("**FirEurisk**")
    ) |> 
  tab_style(
    style = cell_text(weight = "bold"),  # Apply bold text
    locations = cells_body(
      rows = c(33, 34)  # specify the row number
    )
  )

```

### Effect distance to large forest

```{r}
#| label: functions-wui-by-distance-large-forest
# Distance to large forest areas

# Distance variation
max_distances <- c(200, 500, 1000, 2000, 2400)

# WUIs
f_wui <- function(.x, .y){

  veg_perc <- .x

  get_wui_raster(.veg_perc = veg_perc,
                 .veg_dist = .y,
                 .pot_wui = wui_potential,
                 .region = region_v)

}

# Summary table
tbl_summary_distance <- function(.x){

.x |>
  as_tibble() |>
  drop_na() |>
  pivot_longer(everything(),
               names_to = "name",
               values_to = "value") |>
  group_by(name, value) |>
  summarise(count = n()) |>
  ungroup() |>
  group_by(name) |>
  mutate(total = sum(count),
         pct = 100*count/total) |>
  pivot_wider(names_from = "value",
              values_from = c("count", "pct")) |>
  mutate(pct_wui = pct_Interface + pct_Intermix) |>
  select(starts_with("pct")) |>
  relocate(pct_wui, .after = pct_No_wui) |>
  mutate(name = factor(name,
                       levels = c("d200",
                                  "d500",
                                  "d1000",
                                  "d2000",
                                  "d2400"))) |>
  arrange(name) |>
  gt(row_group_as_column = TRUE) |>
  fmt_number(decimals = 1) |>
  cols_label(
    name = "",
    pct_No_wui = md("**No WUI**"),
    pct_Interface = md("**Interface**"),
    pct_Intermix = md("**Intermix**"),
    pct_wui = md("**Total**")
  )  |>
  tab_spanner(
    label = md("**WUI**"),
    columns = c(pct_Interface, pct_Intermix, pct_wui)
  ) |>
  tab_options(
    table.font.size = 9,
    row_group.padding = px(1),
    row_group.font.weight = "bold",
    data_row.padding = px(1)
  )

}

```

#### Effis

```{r}
#| label: fig-wui-effis-vegetations
#| fig-cap: WUI change based on EFFIS fuel map  
#| cache: true

# Raster with distances to large forest (in m)
dist_path <- "data/processed_data/effis_wild_veg_large_dist.tif"
effis_wild_veg_large_dist <- rast(dist_path)

# Withing the max. distance
effis_dist_variation<- map2(effis_wild_veg_large_dist, max_distances, get_dist_max)
names(effis_dist_variation) <- c("d200", "d500", "d1000", "d2000", "d2400")

# Wui vs distance
wui_effis <- map2(effis_wv_perc,
                  effis_dist_variation,
                  f_wui)
wui_effis <- rast(wui_effis)
names(wui_effis) <- c("d200", "d500", "d1000", "d2000", "d2400")
set.values(wui_effis)
terra::saveRDS(wui_effis, "data/processed_data/wui_effis.rsd")

# Plot
panel(wui_effis, axes = FALSE)

```

```{r}
#| label: wui-effis-dist

wui_effis <- terra::readRDS("data/processed_data/wui_effis.rsd")

# Data by country
wui_effis_no <- mask(wui_effis, no_v)
wui_effis_se <- mask(wui_effis, se_v)

```

```{r}
#| label: tbl-wui-effis-dist
#| tbl-cap: Changes in WUI percentages of WUI by country (EFFIS)
#| tbl-subcap: 
#|  - Norway
#|  - Sweden
#| layout-ncol: 2

tbl_summary_distance(wui_effis_no)
tbl_summary_distance(wui_effis_se)

```

#### Corine

```{r}
#| label: fig-wui-corine-vegetations
#| fig-cap: WUI change based on CORINE vegetation map  
#| cache: true

# Raster with distances to large forest (in m)
dist_path <- "data/processed_data/corine_wild_veg_large_dist.tif"
corine_wild_veg_large_dist <- rast(dist_path)

# Withing the max. distance
corine_dist_variation<- map2(corine_wild_veg_large_dist, max_distances, get_dist_max)
names(corine_dist_variation) <- c("d200", "d500", "d1000", "d2000", "d2400")

# Wui vs distance
wui_corine <- map2(corine_wv_perc,
                   corine_dist_variation,
                   f_wui)
wui_corine <- rast(wui_corine)
names(wui_corine) <- c("d200", "d500", "d1000", "d2000", "d2400")
terra::saveRDS(wui_corine, "data/processed_data/wui_corine.rsd")

# Plot
panel(wui_corine, axes = FALSE)

```

```{r}
#| label: wui-corine-dist

wui_corine <- terra::readRDS("data/processed_data/wui_corine.rsd")

# Data by country
wui_corine_no <- mask(wui_corine, no_v)
wui_corine_se <- mask(wui_corine, se_v)

```

```{r}
#| label: tbl-wui-corine-dist
#| tbl-cap: Changes in WUI percentages of WUI by country (CORINE)
#| tbl-subcap: 
#|  - Norway
#|  - Sweden
#| layout-ncol: 2
#| cache: true

tbl_summary_distance(wui_corine_no)
tbl_summary_distance(wui_corine_se)

```

#### National datasets

```{r}
#| label: fig-wui-national-vegetations-maps
#| fig-cap: WUI change based on national vegetation maps  
#| cache: true

# Norway ----
dist_path <- "data/processed_data/sr16_wild_veg_large_dist.tif"
sr16_wild_veg_large_dist <- rast(dist_path)
sr16_dist_variation<- map2(sr16_wild_veg_large_dist, max_distances, get_dist_max)
names(sr16_dist_variation) <- c("d200", "d500", "d1000", "d2000", "d2400")
wui_sr16 <- map2(sr16_wv_perc, sr16_dist_variation, f_wui)
wui_sr16 <- rast(wui_sr16)
names(wui_sr16) <- c("d200", "d500", "d1000", "d2000", "d2400")
terra::saveRDS(wui_sr16, "data/processed_data/wui_sr16.rsd")

# Sweden ----
dist_path <- "data/processed_data/nmd_wild_veg_large_dist.tif"
nmd_wild_veg_large_dist <- rast(dist_path)
nmd_dist_variation<- map2(nmd_wild_veg_large_dist, max_distances, get_dist_max)
names(nmd_dist_variation) <- c("d200", "d500", "d1000", "d2000", "d2400")
wui_nmd <- map2(nmd_wv_perc, nmd_dist_variation, f_wui)
wui_nmd <- rast(wui_nmd)
names(wui_nmd) <- c("d200", "d500", "d1000", "d2000", "d2400")
terra::saveRDS(wui_nmd, "data/processed_data/wui_nmd.rsd")

# Merge datasets ----
wui_national <- mosaic(wui_sr16, wui_nmd, fun = "max") |> 
  add_colours_wui()

# Add names
names(wui_national) <- c("d200", "d500", "d1000", "d2000", "d2400")
set.values(wui_national)
# Export
terra::saveRDS(wui_national, "data/processed_data/wui_national.rsd")

# Plot region  
panel(wui_national, axes = FALSE)

```

```{r}
#| label: wui-national-dist
#| tbl-cap: Changes in WUI percentages of WUI by country (national datasets - SR16 & NMD)

wui_national <- terra::readRDS("data/processed_data/wui_national.rsd")

# Data by country
wui_national_no <- mask(wui_national, no_v)
wui_national_se <- mask(wui_national, se_v)

```

```{r}
#| label: tbl-wui-national-dist
#| tbl-cap: Changes in WUI percentages of WUI by country (national datasets - SR16 & NMD)
#| tbl-subcap: 
#|  - Norway
#|  - Sweden
#| layout-ncol: 2
#| cache: true

tbl_summary_distance(wui_national_no)
tbl_summary_distance(wui_national_se)

```

#### FirEUrisk

```{r}
#| label: fig-wui-firEUrisk-vegetations
#| fig-cap: WUI change based on firEUrisk fuel map  
#| cache: true

# Raster with distances to large forest (in m)
dist_path <- "data/processed_data/firEUrisk_wild_veg_large_dist.tif"
firEUrisk_wild_veg_large_dist <- rast(dist_path)

# Withing the max. distance
firEUrisk_dist_variation<- map2(firEUrisk_wild_veg_large_dist, max_distances, get_dist_max)
names(firEUrisk_dist_variation) <- c("d200", "d500", "d1000", "d2000", "d2400")

# Wui vs distance
wui_firEUrisk <- map2(firEUrisk_wv_perc,
                   firEUrisk_dist_variation,
                   f_wui)
wui_firEUrisk <- rast(wui_firEUrisk)
names(wui_firEUrisk) <- c("d200", "d500", "d1000", "d2000", "d2400")
terra::saveRDS(wui_firEUrisk, "data/processed_data/wui_firEUrisk.rsd")

# Plot
panel(wui_firEUrisk, axes = FALSE)

```

```{r}
#| label: wui-firEUrisk-dist

wui_firEUrisk <- terra::readRDS("data/processed_data/wui_firEUrisk.rsd")

# Data by country
wui_firEUrisk_no <- mask(wui_firEUrisk, no_v)
wui_firEUrisk_se <- mask(wui_firEUrisk, se_v)

```

```{r}
#| label: tbl-wui-firEUrisk-dist
#| tbl-cap: Changes in WUI percentages of WUI by country (firEUrisk)
#| tbl-subcap: 
#|  - Norway
#|  - Sweden
#| layout-ncol: 2
#| cache: true

tbl_summary_distance(wui_firEUrisk_no)
tbl_summary_distance(wui_firEUrisk_se)

```

Aggregate all maps in one figure (Note: data are re-sampled to 500859 cells in the plots)

```{r}
#| label: fig-wui-vegetation-type-distance
#| fig-cap: WUI by vegetation type map and threshold distance
#| fig-height: 15
#| cache: true

# Plot
p_effis <- ggplot() + 
  geom_spatraster(data = wui_effis, maxcell = 2000000) +
  facet_wrap(~lyr, ncol = 1) +
  labs(title = "Effis") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))

p_corine <- ggplot() + 
  geom_spatraster(data = wui_corine, maxcell = 2000000) +
  facet_wrap(~lyr, ncol = 1) +
  labs(title = "Corine") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))

p_national <- ggplot() + 
  geom_spatraster(data = wui_national, maxcell = 2000000) +
  facet_wrap(~lyr, ncol = 1) +
  labs(title = "National") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))

p_firEUrisk <- ggplot() + 
  geom_spatraster(data = wui_firEUrisk, maxcell = 2000000) +
  facet_wrap(~lyr, ncol = 1) +
  labs(title = "firEUrisk") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))

p_effis + p_corine + p_national + p_firEUrisk +
  plot_layout(guides = "collect",
              ncol = 4)

```

Plot variation in the percentage of WUI with the distance to large forest areas by each vegetation/fuel map.

```{r}
#| label: fig-wui-distance
#| fig-cap: Change in the percentage of WUI with the theshold distance to large forest areas

get_tbl_pct <- function(.x) {

  t <- .x |>
    as_tibble() |> 
    drop_na() |> 
    pivot_longer(everything(),
                 names_to = "distance",
                 values_to = "value") |>
    group_by(distance, value) |>
    summarise(count = n()) |>
    ungroup() |> 
    group_by(distance) |>
    mutate(total = sum(count),
           pct = 100*count/total) |> 
    ungroup() |> 
    pivot_wider(names_from = "value",
                values_from = c("count", "pct")) |> 
    select(distance, starts_with("pct")) |> 
    mutate(distance = gsub("d", "", distance),
           distance = as.numeric(distance)) |> 
    arrange(distance)

  return(t)

}

# Combine rasters
raster_list <- list(wui_effis_no,
                    wui_corine_no,
                    wui_national_no,
                    wui_firEUrisk_no,
                    wui_effis_se,
                    wui_corine_se,
                    wui_national_se,
                    wui_firEUrisk_se)

# Tables with percentages
dist_list <- raster_list |> purrr::map(get_tbl_pct)

# Add names to de list
names(dist_list) <- c("Norway_effis",
                      "Norway_corine",
                      "Norway_national",
                      "Norway_firEUrisk",
                      "Sweden_effis",
                      "Sweden_corine",
                      "Sweden_national",
                      "Sweden_firEUrisk")

# Merge with names
dist_tbl <- bind_rows(dist_list, .id = "Ticker") |> 
  separate(Ticker, c("country", "vegetation"))

# add value for 2400 m 
get_tbl_pct_2 <- function(.x){

t <- .x |>
  as_tibble() |> 
  drop_na() |> 
  pivot_longer(everything(),
               names_to = "vegetation",
               values_to = "value") |>
  group_by(vegetation, value) |>
  summarise(count = n()) |>
  ungroup() |> 
  group_by(vegetation) |>
  mutate(total = sum(count),
         pct = 100*count/total) |> 
  ungroup() |> 
  pivot_wider(names_from = "value",
              values_from = c("count", "pct")) |> 
  select(vegetation, starts_with("pct"))

return(t)

}

# Plot
ggplot(data = dist_tbl,
       aes(x = distance,
           y = pct_Interface,
           col = vegetation)) +
  geom_point() +
  geom_line() +
  facet_grid(~country) +
  labs(x = "Distance [m]",
       y = "Perc. [%]"
       # title = "Percentage of WUI Interface vs. distance to forest"
       ) +
  theme_bw() +
  theme(strip.text.x = element_text(size = 12, face = "italic"))

```

### Effect threshold building density

In the following figure we can see the distribution of buildings density (#/$km^2$). The number of buildings in the buffer area is then calculated multiplying the building density by the area of the buffer (i.e., minimum number of buildings (Nmin) = density threshold \[#/$km^2$\] \* Buffer area \[$km^2$\]). In red is the standard threshold for classifying an area as possible WUI, then we have used half of this value (i.e., 3.08 #/$km^2$), representing approximately 25% of the data. Finally, we have also tested with 1 #/$km^2$, which is equivalent to classified any area with buildings as potential WUI.

```{r}
#| label: effect-threshold-build-density

build_focal_no <- mask(build_focal, no_v)

build_focal_se <- mask(build_focal, se_v)

# Get density table

get_d_counts <- function(.x) {

  build_focal |> 
    mask(.x) |> 
    as_tibble() |> 
    drop_na() |> 
    mutate(d_km2 = focal_sum / area_buffer)

}

df_counts <- list(no_v, se_v) |> 
  map(get_d_counts) 
names(df_counts) <- c("Norway", "Sweden")

df_counts_bind <- bind_rows(df_counts, .id = "country")

# Create tables with summary statistics to add to the plot
get_tbl_quantiles <- function(.x) {

  tbl <- .x |> 
    select(d_km2) |>
    quantile(prob = seq(0, 1, 0.25), na.rm = T) |>
    round(2) |> 
    as_tibble() |> 
    rename(d = value) |> 
    mutate(Quantile = c("Min", "Q1", "Median", "Q3", "Max"),
           Quantile = forcats::fct_reorder(Quantile, d)) |> 
    select(Quantile, d)

  tbl_quantiles <- tibble(x = 2800,
                          y = 4.0e+5, 
                          table = list(tbl))

  return(tbl_quantiles)

}

tbl_quantiles <- df_counts |>
  map(get_tbl_quantiles) |> 
  bind_rows(.id = "country")

# Plot
ggplot(df_counts_bind, aes(x = d_km2)) + 
  geom_histogram(bins = 50, color = "darkblue", fill = "lightblue") +
  facet_grid(~country) +
  scale_x_log10() +
  geom_vline(xintercept = 6.172,
             linetype = "dashed",
             color = "red",
             linewidth = 1.3) +
  annotate("text", x = 8, y = 420000,
           label = "6.172", colour = "red", parse = TRUE) + 
  geom_vline(xintercept = 3.08,
             linetype = "dashed",
             color = "orange",
             linewidth = 1.3) +
  annotate("text", x = 4, y = 420000,
           label = "3.08", colour = "orange", parse = TRUE) + 
  geom_vline(xintercept = 1.00,
             linetype = "dashed", 
             color = "blue",
             linewidth = 1.3) +
  annotate("text", x = 1.1, y = 420000,
           label = "1", colour = "blue", parse = TRUE) + 
  labs(title = "Building density",
       x = "#/km2",
       y = "") +
  theme_bw() +
  theme(legend.key = element_rect(fill = NA),
        legend.background = element_rect(color = "gray", linetype = "solid"),
        legend.position = c(.95, .95),
        legend.justification = c("right", "top"),
        legend.box.just = "right",
        legend.margin = margin(6, 6, 6, 6),
        strip.text.x = element_text(size = 12, face = "italic")) +
  # Add tabble with summary statistics
  ggpp::geom_table(data = tbl_quantiles,
                   aes(x = x, y = y, label = table),
                   table.theme = ggpp::ttheme_gtplain(base_size = 12))

```

Potential WUI based on different building density:

#### Building density 3.08 #/km2

```{r}

# Minimum number of buildings (N) in the buffer area ----
# WUI if building density (d) > 6.17 [#/km^2]
# N = d [#/km^2] * A [km^2]
# A = pi * radio^2

N_min_3 <- round(3.08 * area_buffer, 3)

wui_potential_3 <- build_focal
wui_potential_3[wui_potential_3 <= N_min_3] <- NA
wui_potential_3[wui_potential_3 >  N_min_3] <- 1

plot(region_r, col = "lightgrey", axes = F, legend = F)
plot(wui_potential_3, axes = F, legend = F, add = T)

```

```{r}
#| label: calculate-wui3-region
#| cache: true

# EFFIS
effis_wui_3 <- get_wui_raster(.veg_perc = effis_wv_perc,
                              .veg_dist = effis_wild_veg_large_dist_2400,
                              .pot_wui = wui_potential_3)

# CORINE
corine_wui_3 <- get_wui_raster(.veg_perc = corine_wv_perc,
                               .veg_dist = corine_wild_veg_large_dist_2400,
                               .pot_wui = wui_potential_3)

# National maps
sr16_wui_3 <- get_wui_raster(.veg_perc = sr16_wv_perc,
                             .veg_dist = sr16_wild_veg_large_dist_2400,
                             .pot_wui = wui_potential_3,
                             .region = no_v)

nrm_wui_3 <- get_wui_raster(.veg_perc = nmd_wv_perc,
                            .veg_dist = nmd_wild_veg_large_dist_2400,
                            .pot_wui = wui_potential_3,
                            .region = se_v)

national_wui_3 <- mosaic(sr16_wui_3, nrm_wui_3, fun = "max") |> 
  add_colours_wui()

# firEUrisk
firEUrisk_wui_3 <- get_wui_raster(.veg_perc = firEUrisk_wv_perc,
                                  .veg_dist = firEUrisk_wild_veg_large_dist_2400,
                                  .pot_wui = wui_potential_3)

# Put all rasters in one 
wui_r_3 <- c(effis_wui_3, corine_wui_3, national_wui_3, firEUrisk_wui_3)
names(wui_r_3) <- c("effis", "corine", "national", "firEUrisk")

terra::saveRDS(wui_r_3, "data/processed_data/wui_r_3.rds")

```

Map of WUI areas (@fig-wui3-maps-region; @tbl-pct-wui3-region).

```{r}
#| label: read-wui3-raster

wui_r_3 <- terra::readRDS("data/processed_data/wui_r_3.rds")

```

```{r}
#| label: fig-wui3-maps-region
#| fig-cap: "WUI maps of Norway based on different fuel/vegetation maps (resolution of 250 x 250 m)"
#| fig-subcap: 
#|   - "EFFIS"
#|   - "CORINE"
#|   - "FirEUrisk"
#|   - "National databases"
#| layout-ncol: 2
#| fig-width: 4
#| fig-format: png
#| cache: true

f_plot <- function(.x){ plot(.x, 
                             axes = F, 
                             mar = c(0, 0, 0, 0),
                             legend = "topleft")  }

f_plot(wui_r_3$effis)
f_plot(wui_r_3$corine)
f_plot(wui_r_3$firEUrisk)
f_plot(wui_r_3$national)

```

```{r}
#| label: tbl-pct-wui3-region
#| tbl-cap: Percentages of WUI by vegetation
#| tbl-subcap: 
#|  - Norway
#|  - Sweden
#| layout-ncol: 2

# Summary WUI by country
wui_r_3_no <- mask(wui_r_3, no_v)
wui_r_3_se <- mask(wui_r_3, se_v)

# summary table 
tbl_summary_vegetation(wui_r_3_no)
tbl_summary_vegetation(wui_r_3_se)

```

#### Building density 1 #/km2

Equivalent to select any area with building as WUI.

```{r}

# Minimum number of buildings (N) in the buffer area ----
# WUI if building density (d) > 6.17 [#/km^2]
# N = d [#/km^2] * A [km^2]
# A = pi * radio^2

N_min_1 <- round(1.00 * area_buffer, 3)

wui_potential_1 <- build_focal
wui_potential_1[wui_potential_1 <= N_min_1] <- NA
wui_potential_1[wui_potential_1 >  N_min_1] <- 1

plot(region_r, col = "lightgrey", axes = F, legend = F)
plot(wui_potential_1, axes = F, legend = F, add = T)

```

```{r}
#| label: calculate-wui1-region
#| cache: true

# EFFIS
effis_wui_1 <- get_wui_raster(.veg_perc = effis_wv_perc,
                              .veg_dist = effis_wild_veg_large_dist_2400,
                              .pot_wui = wui_potential_1)

# CORINE
corine_wui_1 <- get_wui_raster(.veg_perc = corine_wv_perc,
                               .veg_dist = corine_wild_veg_large_dist_2400,
                               .pot_wui = wui_potential_1)

# National maps
sr16_wui_1 <- get_wui_raster(.veg_perc = sr16_wv_perc,
                             .veg_dist = sr16_wild_veg_large_dist_2400,
                             .pot_wui = wui_potential_1,
                             .region = no_v)

nrm_wui_1 <- get_wui_raster(.veg_perc = nmd_wv_perc,
                            .veg_dist = nmd_wild_veg_large_dist_2400,
                            .pot_wui = wui_potential_1,
                            .region = se_v)

national_wui_1 <- mosaic(sr16_wui_1, nrm_wui_1, fun = "max") |> 
  add_colours_wui()

# firEUrisk
firEUrisk_wui_1 <- get_wui_raster(.veg_perc = firEUrisk_wv_perc,
                                  .veg_dist = firEUrisk_wild_veg_large_dist_2400,
                                  .pot_wui = wui_potential_1)

# Put all rasters in one 
wui_r_1 <- c(effis_wui_1, corine_wui_1, national_wui_1, firEUrisk_wui_1)
names(wui_r_1) <- c("effis", "corine", "national", "firEUrisk")

terra::saveRDS(wui_r_1, "data/processed_data/wui_r_1.rds")

```

Map of WUI areas (@fig-wui1-maps-region; @tbl-pct-wui1-region).

```{r}
#| label: read-wui1-raster

wui_r_1 <- terra::readRDS("data/processed_data/wui_r_1.rds")

```

```{r}
#| label: fig-wui1-maps-region
#| fig-cap: "WUI maps of Norway based on different fuel/vegetation maps (resolution of 250 x 250 m)"
#| fig-subcap: 
#|   - "EFFIS"
#|   - "CORINE"
#|   - "FirEUrisk"
#|   - "National databases"
#| layout-ncol: 2
#| fig-width: 4
#| fig-format: png
#| cache: true

f_plot <- function(.x){ plot(.x, 
                             axes = F, 
                             mar = c(0, 0, 0, 0),
                             legend = "topleft")  }

f_plot(wui_r_1$effis)
f_plot(wui_r_1$corine)
f_plot(wui_r_1$firEUrisk)
f_plot(wui_r_1$national)

```

```{r}
#| label: tbl-pct-wui1-region
#| tbl-cap: Percentages of WUI by vegetation
#| tbl-subcap: 
#|  - Norway
#|  - Sweden
#| layout-ncol: 2

# Summary WUI by country
wui_r_1_no <- mask(wui_r_1, no_v)
wui_r_1_se <- mask(wui_r_1, se_v)

# Tables
tbl_summary_vegetation(wui_r_1_no)
tbl_summary_vegetation(wui_r_1_se)

```

Plot all together

```{r}
#| label: fig-wui-threshold
#| fig-height: 15

p_1 <- ggplot() + 
  geom_spatraster(data = wui_r) +
  facet_wrap(~lyr, nrow = 1) +
  labs(title = "d = 1.00 #/km2") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))

p_3.08 <- ggplot() + 
  geom_spatraster(data = wui_r) +
  facet_wrap(~lyr, nrow = 1) +
  labs(title = "d = 3.08 #/km2") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))

p_6.17 <- ggplot() + 
  geom_spatraster(data = wui_r) +
  facet_wrap(~lyr, nrow = 1) +
  labs(title = "d = 6.17 #/km2") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))

# NOTE: geom_spatraster change the resolution of the map
p_1 / p_3.08 / p_6.17 +
  plot_layout(guides = "collect")

```

# Wildfires datasets

We analyse all wildfires that have affected buildings from 2016 to 2022 (i.e., 330). The data have been obtained from XX (Norway) and XXX (Sweden).

```{r}
#| label: wildfires-points

wf_no <- readxl::read_xlsx("data/wui_fires_norway_wgs84.xlsx", skip = 1) |> 
  janitor::clean_names() |> 
  st_as_sf(coords = c("geolokasjon_lengdegrad", "geolokasjon_breddegrad"),
           crs =  4326) |> 
  select(geometry) |> 
  st_transform(crs = 3035) |> 
  mutate(country = "NO")

wf_se <- readxl::read_xlsx("data/wui_fires_sweden_epsg3006.xlsx") |> 
  janitor::clean_names() |> 
  mutate(datetime = lubridate::as_date(datetime)) |> 
  filter(datetime >= lubridate::as_date("2016-01-01")) |> 
  st_as_sf(coords = c("ecoord", "ncoord"),
           crs =  3006) |> 
  select(geometry) |> 
  st_transform(crs = 3035) |> 
  mutate(country = "SE")

wf <- rbind(wf_no, wf_se)

ggplot() +
  geom_sf(data = region_sf, fill = "lightgrey") +
  geom_sf(data = country_sf, fill = NA) +
  geom_sf(data = wf, col = "red", size = 0.4) +
  theme_void()

```

Intersect points with WUI maps

## Effect distance to large forest

Get if the grid to which the wildfire is located was defined as WUI in the respective maps.

```{r}
#| label: tbl-wildfire-wui-map
#| tbl-cap: "Percentage of wildfires in each WUI class"

# Detect wui class for each point (wildfire)
extract_wui <- function(.x) { extract(.x, wf) }

wf_wui <- list(wui_effis, wui_corine, wui_national, wui_firEUrisk) |> 
  map(extract_wui) |> 
  map(as_tibble)
names(wf_wui) <- c("effis", "corine", "national", "firEUrisk")

# Tables with counts of wildfire in each WUI class
f_counts <- function(.x) {
  .x |> 
  pivot_longer(-ID) |> 
  count(name, value) |> 
  mutate(name = factor(name, 
                       levels = c("d200",
                                  "d500",
                                  "d1000",
                                  "d2000",
                                  "d2400"),
                       ordered = T)) |>
  arrange(name) |> 
  pivot_wider(names_from = name,
              values_from = n) |> 
  mutate(across(.cols = starts_with("d"),
                .fns = ~ 100 * (.x / sum(.x)),
                .names = "perc_{.col}"))

}

wf_wui_counts <- wf_wui |>
  map(f_counts) |> 
  # Merge with names
  bind_rows(.id = "vegetation")

# Print table
wf_wui_counts |> 
  # select("vegetation", "value", starts_with("perc")) |> 
  gt(groupname_col = "vegetation",
     row_group_as_column = FALSE) |> 
  fmt_number(columns = starts_with("perc"),
             decimals = 1) |> 
  cols_label(
    vegetation = "",
    value = md("**WUI class**"),
    perc_d200 = md("**d200**"),
    perc_d500 = md("**d500**"),
    perc_d1000 = md("**d1000**"),
    perc_d2000 = md("**d2000**"),
    perc_d2400 = md("**d2400**")
  )  |> 
  cols_label_with(fn = function(x) {
    janitor::make_clean_names(x, case = "title") |>
      tolower() |>
      stringr::str_replace_all("^|$", "**") |>
      md()
  }) |> 
  tab_spanner(
    label = md("**Percentage**"),
    columns = starts_with("perc")
  ) |> 
  tab_spanner(
    label = md("**Count**"),
    columns = starts_with("d")
  ) |> 
  tab_options(
    table.font.size = 9,
    row_group.background.color = "lightgrey",
    row_group.padding = px(1),
    row_group.font.weight = "bold",
    data_row.padding = px(1)
  )

```

## Effect threshold building density

```{r}
#| label: read-wui-maps-building-density

wui_r_1 <- terra::readRDS("data/processed_data/wui_r_1.rds")
wui_r_3 <- terra::readRDS("data/processed_data/wui_r_3.rds")
wui_r <- terra::readRDS("data/processed_data/wui_r.rds")

```

```{r}
#| label: tbl-wildfire-wui-map-building-density
#| tbl-cap: "Percentage of wildfires in WUI calss based on building density threshold"

# Detect wui class for each point (wildfire)
extract_wui <- function(.x) { extract(.x, wf) }

wf_wui_bd <- list(wui_r_1, wui_r_3, wui_r) |> 
  map(extract_wui) |> 
  map(as_tibble)
names(wf_wui_bd) <- c("bd_1", "bd_3", "bd_6")

# Tables with counts og the N. wildfire in each wui class
f_counts <- function(.x) {
  .x |> 
  pivot_longer(-ID) |> 
  count(name, value) |> 
  mutate(name = factor(name, 
                       levels = c("effis",
                                  "corine",
                                  "national",
                                  "firEUrisk"),
                       ordered = T)) |>
  arrange(name) |> 
  pivot_wider(names_from = name,
              values_from = n) |> 
  mutate(across(.cols = -value,
                .fns = ~ 100 * (.x / sum(.x)),
                .names = "perc_{.col}"))

}

wf_wui_bd_counts <- wf_wui_bd |>
  map(f_counts) |> 
  # Merge with names
  bind_rows(.id = "threshold")

# Print table
wf_wui_bd_counts |> 
  mutate(definition = case_when(
    threshold == "bd_1" ~ "threshold = 1.00 #/km2",
    threshold == "bd_3" ~ "threshold = 3.08 #/km2",
    threshold == "bd_6" ~ "threshold = 6.17 #/km2"
  )) |> 
  select(-threshold) |> 
  gt(groupname_col = "definition",
     row_group_as_column = FALSE) |> 
  fmt_number(columns = starts_with("perc"),
             decimals = 1) |> 
  cols_label(
    definition = "",
    value = md("**WUI class**"),
    effis = md("**effis**"),
    corine = md("**corine**"),
    national = md("**national**"),
    firEUrisk = md("**firEUrisk**"),
    perc_effis = md("**effis**"),
    perc_corine = md("**corine**"),
    perc_national = md("**national**"),
    perc_firEUrisk = md("**firEUrisk**")
  )  |> 
  tab_spanner(
    label = md("**Percentage**"), columns = starts_with("perc")
  ) |> 
  tab_spanner(
    label = md("**Count**"), columns = c(effis, corine, national, firEUrisk)) |> 
  tab_options(
    table.font.size = 9,
    row_group.background.color = "lightgrey",
    row_group.padding = px(1),
    row_group.font.weight = "bold",
    data_row.padding = px(1)
  )

```